---
title: "Clean Demographic"
author: "Yan Zhu"
date: "`r lubridate::today()`"
format: 
  html:
    timeout: 180
    embed-resources: true
    toc: true 
    toc_depth: 4
editor_options: 
  chunk_output_type: inline
---

## Introduction
In this file, I will clean and preprocess Demographic data from Baseline data of ESM study. The goal is to prepare the data for analysis.

**Note:** The data file is from the CSV file which generated from Data Anonymization.Rmd

Check and view all the variable of Demographic data.

## Set up
### load packages
```{r set up, echo=FALSE}
library(tidyverse)
library(dplyr)
library(skimr)
library(janitor)
library(skimr)
source("/Users/yan/Desktop/sample_code/functions.R")
```

### read data
```{r read data}
# dis2, dis_inj, dis_combined, dis_mult, rel, smu (In old code book, but not in the data)
# ADD! I will create a new Code Book which could be found in repo/sample_code

demog <- read_csv("/Users/yan/Desktop/sample_code/ESM/1_anonymization/anonymous_baseline.csv") %>% 
  select(random_id, age, res, sch_yr, race, mult, race_txt, eth, dis1, dis_dev,
         dis_lrn, dis_sen, dis_mob, dis_psy, dis_unl, dis_txt, dis_ill,dis_dev,
         gend, trans, sx_ori,, dem12, f_aid, p_grnt, job, f_need, fgen) %>% glimpse(show_col_types = FALSE)
```

## Variable Check

### Age
Variable code as "age"
Survey Question: "How old are you?"
```{r}
ggplot(demog, aes(x = age)) +
  geom_bar(fill = "skyblue", color = "black") +
  geom_text(stat = "count", aes(label = after_stat(count)), vjust = -0.5) +
  scale_x_continuous(breaks = seq(min(demog$age), max(demog$age), by = 1)) +
  labs(x = "Age",
       y = "Frequency",
       title = "Histogram of Age",
       subtitle = paste("Sample Size:", nrow(demog))) +
  theme(plot.subtitle = element_text(size = 10)) +
  theme_bw()
```

### Residence
Variable code as "res"
Where do you currently live?
```{r}
table(demog$res)
```

**Note:** 1 = On campus in a dorm , 2 = On campus in an apartment, 3 = Other on campus arrangement, 4 = Off campus apartment, 5 = Off campus house, 6 = Off campus with parents, 7 = On campus fraternity/soroity house, 8 = Off campus fraternity/sorority house,  9 = Other

The numbers in the res column are replaced by corresponding labels, making the data more readable.
```{r}
demog$res <- ifelse(demog$res == 1, "On campus in a dorm", demog$res)
demog$res <- ifelse(demog$res == 2, "On campus in an apartment", demog$res)
demog$res <- ifelse(demog$res == 4, "Off campus apartment", demog$res)
demog$res <- ifelse(demog$res == 5, "Off campus with parents", demog$res)
demog$res <- ifelse(demog$res == 6, "Off campus with parents", demog$res)
demog$res <- ifelse(demog$res == 8, "Off campus fraternity/sorority house", demog$res)
prop_tab(demog, res)
```

### School Tear
Variable code as "sch_yr"
What is your year in school?
```{r}
prop_tab(demog, sch_yr)
```

### Race
Variable code as "race"
Recognizing that race is complex, which of the following classifications of race best describe how you identify?

```{r}
# Checking NA
sum(is.na(demog$race))
```

Most of the variables having NA values is reasonable because there were no applicable options to describe our participants. However, Race has 2 missing data points which seem unexpected if "mult" and "race_txt" are also NA because they are open text-entry responses. Let's check those two rows.
```{r}
# check the rows with NA in Race
demog[is.na(demog$race), ]
```

*There were no information in race_txt and mult(Open text-entry responses from individuals identifying as multiracial). I would label it with "Unknown" string*

```{r}
# Change NA to string.
demog$race <- ifelse(is.na(demog$race), "Unknown", demog$race)
prop_tab(demog, race)
```

More data cleaning on race! Text-entry is also an unexpected answer.
```{r}
# make sure if "Text-entry" with other information about race in other columns
demog[demog$race == "Text-entry", ]
```

*Nope, I will replace "Unknown" instead of "Text-entry"*
```{r}
demog$race <- ifelse(demog$race == "Text-entry", "Unknown", demog$race)
prop_tab(demog, race)
```

```{r}
demog[!is.na(demog$race_txt), ]
```

```{r}
demog[!is.na(demog$mult), ]
```

### Ethnicity
Do you identify as Hispanic or Latino/Latina or Latinx?
```{r}
prop_tab(demog, eth)
# 1 = Yes, 0 = No, 3 = prefer not to respond
```

### Disability
Do you consider yourself to have a disability or be disabled? (e.g., intellectual disability, learning disability, sensory disability, mobility disability, mental illness, etc.)
```{r}
prop_tab(demog, dis1)
```

```{r}
prop_tab(demog, dis_dev)
```


```{r}
# Select columns containing "dis_" in their column names
dis_columns <- demog %>%
  select(contains("dis_"))

# Count the number of NAs in each column
na_counts <- colSums(is.na(dis_columns))
na_counts
```

Except "dis_txt", the count the number of columns which containing "dis_" should be 221. However, the NA count here seems not right.
```{r}
# List the column names that need to be replaced
cols_to_replace <- c("dis_dev", "dis_lrn", "dis_sen", "dis_mob", "dis_psy", "dis_unl", "dis_ill")

# Use mutate and across to replace
demog <- demog %>%
  mutate(across(all_of(cols_to_replace), ~ifelse(is.na(.), 0, .)))

demog <- demog %>%
  mutate(dis_txt = ifelse(is.na(dis_txt), "None", dis_txt))
```



### Gender
Recognizing that gender is complex, how would you describe your gender?
```{r}
# 1 = Man, 2 = Woman, 3 = Non-binary, 4 = Gender fluid, 5 = Two-Spirit, 6 = Agender, 7 = Pangender, 8 = Another gender identity [Write in]
prop_tab(demog, gend)
```

```{r}
demog$gend <- ifelse(demog$gend == 1, "Male", demog$gend)
demog$gend <- ifelse(demog$gend == 2, "Female", demog$gend)
demog$gend <- ifelse(demog$gend == 4, "Other", demog$gend)
demog$gend <- ifelse(demog$gend == 5, "Other", demog$gend)
prop_tab(demog, gend)
```


Question: Are you transgender?
```{r}
# 1 = Yes, 0 = No
prop_tab(demog, trans)
```

Question: Recognizing that sexual orientation can be fluid, how would you describe your current sexual orientation?
```{r}
prop_tab(demog, sx_ori)
```

```{r}
# 1 = Gay, 2 = Lesbian, 3 = Bisexual, 4 = Pansexual, 5 = Heterosexual/Straight, 6 = Other [Write in]
demog$sx_ori <- recode(demog$sx_ori, 
                       `1` = "Gay",
                       `2` = "Lesbian",
                       `3` = "Bisexual", 
                       `4` = "Pansexual",
                       `5` = "Heterosexual",
                       `6` = "Other")

prop_tab(demog, sx_ori)
```


```{r}
glimpse(demog$dem12)
```


Question: Do you participate in any of the following activities (check all that apply):
```{r}
# These question is a multiple choice question, and each number represent an activity option.
## Select all that apply: 1=Athletics; 2=Club sports; 3=Greek life; 4=Registered student organizations; 5=Volunteer work; 6=Associated students of Madison; 7=Work-study; 8=Research Involvement; 9=Religious-affiliated student organizations
demog$dem12 <- sapply(strsplit(as.character(demog$dem12), ""), function(x) paste(x, collapse = ","))
glimpse(demog$dem12)
```


Count number of acvitity were select in question dem12.
```{r}
demog$num_activity <- str_count(demog$dem12, "\\d+")
range(demog$num_activity)
```

Do you receive financial aid?
```{r}
# 1 = Yes, 0 = No
prop_tab(demog, f_aid)
```

Have you ever received a Pell grant?
```{r}
# 1 = Yes, 0 = No
prop_tab(demog, p_grnt)
```

Do you work for pay during the school year?
```{r}
# 1 = Yes, 0 = No
prop_tab(demog, job)
```

Since you have been a student at UW-Madison, how often have you had difficulty paying for basic necessities like food, clothing, or rent?
```{r}
# 1 = Never, 2 = Rarely, 3 = Sometimes, 4 = Often, 5 = All the time
prop_tab(demog,f_need)
```

Would you consider yourself a first-gen college student (meaning your parents never completed a 4-year college or university degree)? 
```{r}
# 1 = Yes, 0 = No, 3 = Not sure
prop_tab(demog, fgen)
```

```{r}
demog <- demog %>% select(random_id, age, res, sch_yr, race, 
                          mult, race_txt, eth, dis1, dis_dev,
                          dis_lrn, dis_sen, dis_mob, dis_psy, dis_unl, 
                          dis_txt, dis_ill, gend, trans, sx_ori,
                          dem12, num_activity, everything())
glimpse(demog)
```


## Write CSV
```{r}
write.csv(demog, "/Users/yan/Desktop/sample_code/ESM/2_demograpgic/demog_data.csv", row.names = FALSE)
```







