---
title: "Partnered Sexual Behavior"
author: "Yan Zhu"
date: "`r lubridate::today()`"
format: 
  html:
    timeout: 180
    embed-resources: true
    toc: true 
    toc_depth: 4
editor_options: 
  chunk_output_type: inline
---
## Introduction

Check out the the survey about Partnered Sexual Behavior.

These survey were answered because participants answered **YES** in Question: *Since the last survey, have you hooked up, engaged in any sexual activity, or tried to engage in any sexual activity with another person (including flirting, kissing, touching, or intercourse)?*

Aims of this Script:

1. Check and understanding this dataset.

2. Clean it.


## Set Up
```{r setup, echo=TRUE, message=FALSE}
library(tidyverse)
library(janitor)
library(skimr)
library(psych)
library(knitr)
source("/Users/yan/Desktop/sample_code/functions.R")
```

Read data
```{r read data, message=FALSE}
daily <- read_csv("/Users/yan/Desktop/sample_code/ESM/1_anonymization/anonymous_daily.csv") %>% clean_names()
```

```{r}
psb1 <- daily %>%
  filter(psb1 == 1) %>%
  select(random_id, start_date, end_date, finished, psb2, cub1, cub2, cub3, cub4, cub5, cub6, psb3, psb4, psb5, psb5a, psb6, psb6a, str_p, str_p_txt, psb7, psb8, psb9, psb7a, psb8a, psb9a, psb10, psb10a, psb11, psb11a, med1_p, med2_p, med3_p, med4_p, aro_pf, pls_pf, con_pf, ver_pc, enon_pc, inon_pc, prp_v, prp_p, prp_sf, prp_in, vic_v, vic_p, vic_sf, vic_in, exp_p, st_p, pa1_p, na1_p, na2_p, na3_p, na4_p, na5_p, pa2_p, er1_p, er2_p, er3_p, er4_p) %>%
  glimpse()
```

**Question:** How many Partnered Sexual Behavior survey were recorded? (finished?)
```{r}
cat("We have", nrow(psb1), "records.")
```

```{r}
finished <- psb1[psb1$finished == 1, ]
unfinished <- psb1[psb1$finished == 0, ]

cat("There are", nrow(finished), "records; and", nrow(unfinished), "unfinished records.")
```

Let's remove column contains only a single unique value.
```{r}
# Checks each column to determine if it contains only a single unique value
single_valued_columns <- sapply(finished, function(col) length(unique(col)) == 1)
cat("Variables with only one unique value:\n", names(finished[single_valued_columns]))
```

```{r}
glimpse(finished$str_p_txt)
```

```{r}
# All NA is not helpful, so remove this column.
finished <- finished %>% select(-str_p_txt)
```

## Emotion Regulation
I will focuse on participant with Emotion Regulation scores.
```{r}
# remove the rows without Emotion Regulation score.
finished <- finished[!is.na(finished$er1_p) & !is.na(finished$er1_p) & !is.na(finished$er1_p) &  !is.na(finished$er1_p), ]

cat("We have", nrow(finished), "records now.")
```

er1_p: "Thinking about your strongest emotion, how did you respond to the feeling?  I tried to get rid of negative thoughts, feelings, or sensations"

- This could be seen as a form of emotional suppression, which is often considered maladaptive. So, if "yes" means the respondent engaged in this behavior, it does not require reverse coding.

er2_p: "Thinking about your strongest emotion, how did you respond to the feeling?  I couldn’t stop thinking about my feelings"

- This indicates rumination, which is also considered maladaptive. A "yes" response here should also not be reverse-coded.

er3_p: "Thinking about your strongest emotion, how did you respond to the feeling?  I avoided expressing my feelings"

- Avoidance of expressing feelings is another maladaptive strategy. A "yes" response here would not need reverse coding.

er4_p: "Thinking about your strongest emotion, how did you respond to the feeling?  I changed the way I thought about what caused my feelings"

- This is potentially an adaptive strategy, often related to cognitive reappraisal. If "yes" indicates a positive, adaptive behavior, this item require reverse coding to align with the others, where "yes" indicates a negative behavior.

```{r}
# Get the total score of emotion regulation.
finished$er_sum <- finished$er1_p + finished$er2_p + finished$er3_p - finished$er4_p
prop_tab(finished, er_sum)
```


```{r}
er_p <- finished %>% select(random_id, start_date, er1_p, er2_p, er3_p, er4_p, er_sum)
```

```{r}
er_p$start_date <- as.Date(er_p$start_date)

id_counts <- er_p %>%
  count(random_id, name = "count")
```
















